The Power of Two Random Choices

### 2014-03-25 presentation/presentation.tex
some minor fixes+#

\begin{frame}
\frametitle{Single choice}
\framesubtitle{Upper bound for the maximum load}
\begin{theorem}[Raab and Steger, 1998]
When using the single choice strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\load = 
	\begin{cases}
    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases}
\end{align*}
\end{theorem}
\end{frame}

@@

\pause
\bigskip
We can distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
\begin{itemize}
\item Uniformly and independently
\item Non-unfiromly and independently
\item Non-uniformly and dependently
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Greedy}
\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be (w.h.p.~and for all $m$): 
\begin{align*}
\load = \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
\end{align*}
\end{theorem}
\pause
\medskip
Compared to the \alert{single choice} strategy, with 
\begin{align*}
\load = 
	\begin{cases}
    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad ,
\end{align*}
\alert{Greedy} with $d=2$:
\begin{itemize}
\item achieves an \alert{exponential decrease} of the additive gap $\delta$ and
\item $\delta$ does \alert{not depend on} the number of balls \alert{$m$} anymore.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Always-go-left}
\begin{theorem}[V\"ocking, 2003]
When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be  (w.h.p.~and for all $m$): 
\begin{align*}
\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
\end{align*}
\end{theorem}
\todo{$\Phi_d$ is undefined. I think a full definition only makes sense if we present the proof for Always-go-left.}

\bigskip
The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.
\end{frame}

@@

\begin{description}
\item[\alert{Sequential:}] The balls arrive one after another.
\item[\alert{Infinite:}] Ball placements and deletions are given for an infinite time span.
\item[\alert{Dynamic:}] At each time step $t$, one ball is removed from the bins.
\begin{itemize}
\item An oblivious adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$.
\item $d_t$/$a_t$: ball to delete/add at time step t. \\
\item $\sigma$ is fixed in advance.
\end{itemize}
\end{description}
\end{frame}

\begin{frame}
\frametitle{Proof of either Greedy or Always-go-left}
\end{frame}

@@

%\againframe{references}

\begin{frame}[allowframebreaks, label=references]
\frametitle<presentation>{References}    
\bibliographystyle{acm}
\nocite{*}
\setbeamertemplate{bibliography item}[text]
\bibliography{../sources}
\end{frame}

>>>>

\begin{frame}
\frametitle{Single choice}
\framesubtitle{Upper bound for the maximum load}
\begin{theorem}[Raab and Steger, 1998]
When using the single choice strategy to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\load = 
	\begin{cases}
    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad .
\end{align*}
\end{theorem}
\end{frame}

@@

\pause
\bigskip
We can distinguish placement algorithms by the way the $d$ \alert{bins are sampled}:
\begin{itemize}
\item Uniformly and independently
\item Non-uniformly and independently
\item Non-uniformly and dependently
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Greedy}
\begin{theorem}[Azar, Broder, Karlin and Upfal, 1999]
When using the \alert{Greedy} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\load = \frac{m}{n} + \frac{\ln \ln n}{\ln d} + O(1) \quad .
\end{align*}
\end{theorem}
\pause
\medskip
Compared to the \alert{single choice} strategy, with 
\begin{align*}
\load = 
	\begin{cases}
    \frac{\ln n}{\ln\ln n} + O(1)              & m = n \\
    \frac{m}{n} + \Theta\left(\sqrt{\frac{m\ln n}{n}} \right)              & m \in \omega(n \ln n)
    \end{cases} \quad ,
\end{align*}
\alert{Greedy}
\begin{itemize}
\item achieves an \alert{exponential decrease} of the additive gap $\delta$ and
\item $\delta$ does \alert{not depend on} the number of balls \alert{$m$} anymore.
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{Always-go-left}
\begin{theorem}[V\"ocking, 2003]
When using the \alert{Always-go-left} algorithm to distribute $m$ balls to $n$ bins, the maximum load will be, w.h.p.,
\begin{align*}
\load = \frac{m}{n} + \frac{\ln \ln n}{d \ln \Phi_d} + O(1) \quad .
\end{align*}
\end{theorem}
\todo{$\Phi_d$ is undefined. I think a full definition only makes sense if we present the proof for Always-go-left.}

\pause
\bigskip
The combination of \alert{asymmetric tie breaking} and \alert{non-uniform} (and independent) sampling of the $d$ bins is crucial to achieve this upper bound.

\medskip
V\"ocking further proved a matching lower bound for $\load$, i.e., the upper bound is tight.

@@

\begin{description}
\item[\alert{Sequential:}] The balls arrive one after another, one at each time step $t$.
\item[\alert{Infinite:}] Ball placements and deletions are given for an infinite time span.
\item[\alert{Dynamic:}] Balls do not stay in the bins forever.
\begin{itemize}
\item An oblivious adversary provides a sequence $\sigma = \sigma_1 \sigma_2 \dots$, where $\sigma_t = (d_t, a_t)$.
\item $d_t$/$a_t$: ball to delete/add at time step t. \\
\item $\sigma$ is fixed in advance.
\end{itemize}
\end{description}
\end{frame}

\begin{frame}
\frametitle{Proof of either Greedy or Always-go-left}
I suggest to present the Greedy proof:
\begin{itemize}
\item Both proofs have the same structure, but Always-go-left is a lot more technical
\item For Always-go-left, d-ary Fibonacci trees and a generalized golden ratio need to be introduced
\item complete d-ary trees are easier to think and reason about (number of nodes/leaves, the inductive argument used)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Witness trees}
\end{frame}

\begin{frame}
\frametitle{Witness trees}
\begin{alertblock}{Assumption}
All the events represented by the nodes and edges of a witness tree are stochastically independent. 
\end{alertblock}
\end{frame}

@@

%\againframe{references}
%todo: why so few references shown?
\begin{frame}[allowframebreaks, label=references]
\frametitle<presentation>{References}    
\bibliographystyle{acm}
\nocite{*}
\setbeamertemplate{bibliography item}[text]
\bibliography{../sources}
\end{frame}
